%TEX
%
\documentclass[11pt]{CAS_syllabus}
\usepackage{lastpage}
\usepackage[text={6.75in,9.3in}, top=0.8in, left=0.9in]{geometry}

%-------------- begin the document --------------------------------------------
\begin{document}

% Define the header for the Course Syllabus document:
\Heading

\Coursehead[
% The following defines the heading of the document.
instructor      = {Landon Rabern},
phone           = {},
email           = {lrabern@newhaven.edu},
course          = {MATH 3305},
coursecrosslist = {},
coursename      = {Discrete Structures},
revision        = {1.0, 08/18/2013},
semester        = {Fall 2013},
office          = {},
meetingtimes    = {TTh 12:15--13:30},
classroom       = {Maxcy 203},
credits         = {3},
officehours     = {TTh 13:30pm--15:00},
%
%
%
%
% The following are web links used throughout the text of the document.
addanddrop              = {Monday, September, 9},
addanddropurl           = {http://www.newhaven.edu/academics/13324/},
addanddroppolicyurl     = {http://www.newhaven.edu/academics/13324/},
attendanceurl           = {http://www.newhaven.edu/academics/16648/},
integrityurl            = {http://www.newhaven.edu/academics/16246/},
courseworkurl           = {http://www.newhaven.edu/academics/16697/},
accessurl               = {http://www.newhaven.edu/student-life/CampusLife_StudentAffairs/Campus_Access_Services/},
undergraduatecalendarurl= {http://www.newhaven.edu/academics/13324/},
finalexamurl            = {http://www.newhaven.edu/academics/29258/},
importantdatesurl       = {http://www.newhaven.edu/academics/12592/},
gradepolicyurl          = {http://www.newhaven.edu/academics/17041/},
%
% The following are dates used throughout the text of the document.
semesterclasses      = {Fall Semester 2013},
startsemesterclasses = {Thursday, August 23},
endsemesterclasses   = {Monday, December 10},
lastdropdate         = {Monday, September 9},
semesterholidays     = {Labor Day, September 2; Fall Break, Monday/Tuesday October 14--15; Thanksgiving, November 26 -- November 29},
semesterfinalexam    = {Thursday December 12, from 14:00 to 16:00 in Maxcy 203}
]

%...............................................................................

\Title{\Course\ \Coursename}

\Subtopic{Course Description}
Prerequisite: MATH 1108. Methods of proof, the integers, induction, prime numbers, recursive algorithms, greatest common divisors, the Euclidean algorithm, the fundamental theorem of arithmetic, congruences. 3 credits. 


\Subtopic{Required Text}
\textit{A Friendly Introduction to Number Theory}, Joseph H.\ Silverman,
Pearson (2013), ISBN 13: 978-0-321-81619-1.

\Subtopic{Recommended Additional Text}
\textit{Proofs from the Book}, Martin Aigner, Gunter M. Ziegler, 4th ed. (2009).

\Topic{Content}
\Course\ provides an introduction to the integers and their fundamental nature.
In rigorously investigating the integers, students will learn the basic proof techniques used in all of mathematics and increase their mathematical maturity.


%\Subtopic{Topics}
%\begin{compactenum}
%\item
%\end{compactenum}


%STANDARD EXAM POLICY
% SELECT:
% \input{policy-exam2}
% \input{policy-exam3}
\input{policy-exam}
% END EXAM POLICY

\Subtopic{Further Considerations and Rules}
\begin{compactitem}
\item
\textit{Electronic devices}:
The use of phones and electronic devices, except in an emergency or unless
otherwise approved by the instructor,
is disallowed in class. The use of any electronic device, except those
explicitly approved for use during any exam, will result a non-positive grade equal to the number of malus points on that exam. 
If in doubt, please inquire prior to using the electronic device.
\item
\textit{Scheduling conflicts}:
It is the student's responsibility to contact the instructor concerning any
scheduling conflicts which may result in late papers, or other scheduling
conflicts, e.g., an absence for an exam.
\end{compactitem}


\Subtopic{Course Outline Schedule}
\Semester\ classes are from \Startsemesterclasses\ to \Endsemesterclasses.
The last day to drop classes is \Lastdropdate.
Semester holidays are on: \Semesterholidays.
The final exam is scheduled for \Semesterfinalexam.  A schedule of all assignments due will be available on blackboard.

%COPIES OF THE SCHEDULE FOR MW AND TR ARE KEPT IN ./lib.........................
{
\renewcommand{\syllabusfontstable}{\syllabusfontf}
\renewcommand{\tableheader}{\tableheaders}
\input{./schedule.tex}
}



\syllabusfontn
The Undergraduate Academic Calendar is found
at \Urls{\UndergraduatecalendarURL}, and the date for final exam dates can
be found at \Urls{\FinalexamURL}. A summary of important dates can also be
found at \Urls{\ImportantdatesURL}.



%THIS SECTION IS STANDARD POLICY....................................DO NOT ALTER
%\newpage
\input{CASpolicy}
%THIS SECTION IS STANDARD POLICY....................................DO NOT ALTER

\end{document}

