\Topic{Grading}
Students will be evaluated on the basis of three exams (worth $30$ points each) and one project (worth $10$ points). Total points (T) map to grades as follows.

\begin{center}
\begin{tabular}{rcr|l}
\multicolumn{3}{c}{\textbf{T}} & \textbf{grade} \\
\hline
& & &  \\[-0.8em]
    $97$ &-- &  $\infty$ & \hspace*{0.8em} \ec A+ \\
    $93$ &-- &  $96$ & \hspace*{0.8em} \ec A \\
    $90$ &-- &  $92$ & \hspace*{0.8em} \ec A- \\
    $87$ &-- &  $89$ & \hspace*{0.8em} \ec B+ \\
    $83$ &-- &  $86$ & \hspace*{0.8em} \ec B \\
    $80$ &-- &  $82$ & \hspace*{0.8em} \ec B- 
\end{tabular} \hspace*{1.8em}
\begin{tabular}{rcr|l}
\multicolumn{3}{c}{\textbf{T}} & \textbf{grade} \\
\hline
& & &  \\[-0.8em]
    $77$ &-- &  $79$ & \hspace*{0.8em} \ec C+ \\
    $73$ &-- &  $76$ & \hspace*{0.8em} \ec C \\
    $70$ &-- &  $72$ & \hspace*{0.8em} \ec C- \\
    $67$ &-- &  $69$ & \hspace*{0.8em} \ec D+ \\
    $63$ &-- &  $66$ & \hspace*{0.8em} \ec D \\
    $-\infty$ &-- & $62$ & \hspace*{0.8em} \ec F
\end{tabular}
\end{center}

More precisely, exams 1, 2 and 3 are given scores $E_1$, $E_2$, $E_3$, the project is given score $P$ and we let $T = \max\{E_1, E_3\} + \max\{E_2, E_3\} + E_3 + P$.  The $\max\{E_1, E_3\}$ and $\max\{E_2, E_3\}$ terms serve the following purpose: as exam 3 covers all material in the course, a student that does well on it should get a good grade overall.    For example, a student getting $0$ points on exam 1, exam 2 and the project, but getting $30$ points on exam 3 will get $T = \max\{0, 30\} + \max\{0, 30\} + 30 + 0 = 90$ which is an $A-$.

Exams will have bonus problems at the end that require some cleverness to solve.  A correct answer to a bonus problem is worth $+1$ point; an incorrect answer doesn't lose a point, but a correct answer wins a point.  By symmetry, exams will also have malus problems that every student should be able to answer (most often these will just be asking for basic definitions). An incorrect answer to a malus problem is worth $-1$ point; a correct answer doesn't gain a point, but an incorrect answer loses a point.

\Topic{The Project}
There will be a project involving the programming language FRACTRAN (\url{http://en.wikipedia.org/wiki/FRACTRAN}). This is an extremely simple language that uses prime exponents as registers.  More details about the project will come after we have introduced prime numbers and factorization.

\Topic{Homework}
Homework does not count for points; however, since the exam problems will be similar to homework problems, it would be irrational to not work the assigned problems.