\documentclass[12pt]{article}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage[top=1.0in, bottom=1.0in, left=1.0in, right=1.0in]{geometry}

\pagestyle{plain}

\usepackage{tkz-graph}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}
\usepackage[position=bottom]{subfig}
\usepackage{graphicx}


\usepackage{longtable}
\usepackage{array}

\usepackage{sectsty}
\allsectionsfont{\sffamily}

\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}

\makeatletter
\newtheorem*{rep@theorem}{\rep@title}
\newcommand{\newreptheorem}[2]{
\newenvironment{rep#1}[1]{
 \def\rep@title{#2 \ref{##1}}
 \begin{rep@theorem}}
 {\end{rep@theorem}}}
\makeatother

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newreptheorem{thm}{Theorem}
\newtheorem{prop}[thm]{Proposition}
\newreptheorem{prop}{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newreptheorem{lem}{Lemma}
\newtheorem{conjecture}[thm]{Conjecture}
\newreptheorem{conjecture}{Conjecture}
\newtheorem{cor}[thm]{Corollary}
\newreptheorem{cor}{Corollary}
\newtheorem{prob}[thm]{Problem}

\newtheorem*{SmallPotLemma}{Small Pot Lemma}
\newtheorem*{BK}{Borodin-Kostochka Conjecture}
\newtheorem*{BK2}{Borodin-Kostochka Conjecture (restated)}
\newtheorem*{Reed}{Reed's Conjecture}
\newtheorem*{ClassificationOfd0}{Classification of $d_0$-choosable graphs}


\theoremstyle{definition}
\newtheorem{defn}{Definition}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{problem}{Problem}
\newtheorem{example}{Example}
\newtheorem*{question}{Question}
\newtheorem*{observation}{Observation}

\newcommand{\fancy}[1]{\mathcal{#1}}
\newcommand{\C}[1]{\fancy{C}_{#1}}
\newcommand{\IN}{\mathbb{N}}
\newcommand{\IR}{\mathbb{R}}
\newcommand{\G}{\fancy{G}}
\newcommand{\CC}{\fancy{C}}
\newcommand{\D}{\fancy{D}}

\newcommand{\inj}{\hookrightarrow}
\newcommand{\surj}{\twoheadrightarrow}

\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\setb}[3]{\left\{ #1 \in #2 \mid #3 \right\}}
\newcommand{\setbs}[2]{\left\{ #1 \mid #2 \right\}}
\newcommand{\card}[1]{\left|#1\right|}
\newcommand{\size}[1]{\left\Vert#1\right\Vert}
\newcommand{\ceil}[1]{\left\lceil#1\right\rceil}
\newcommand{\floor}[1]{\left\lfloor#1\right\rfloor}
\newcommand{\func}[3]{#1\colon #2 \rightarrow #3}
\newcommand{\funcinj}[3]{#1\colon #2 \inj #3}
\newcommand{\funcsurj}[3]{#1\colon #2 \surj #3}
\newcommand{\irange}[1]{\left[#1\right]}
\newcommand{\join}[2]{#1 \mbox{\hspace{2 pt}$\ast$\hspace{2 pt}} #2}
\newcommand{\djunion}[2]{#1 \mbox{\hspace{2 pt}$+$\hspace{2 pt}} #2}
\newcommand{\parens}[1]{\left( #1 \right)}
\newcommand{\brackets}[1]{\left[ #1 \right]}
\newcommand{\DefinedAs}{\mathrel{\mathop:}=}

\def\adj{\leftrightarrow}
\def\nonadj{\not\!\leftrightarrow}

\def\D{\fancy{D}}
\def\C{\fancy{C}}

\newcommand{\bigclique}[1]{\frac{2}{3}\Delta(#1) + 5}
\newcommand{\bigcliqueraw}{\frac{2}{3}\Delta + 5}
\newcommand{\cliqueparts}{\frac{2}{3}\Delta}


\title{Coloring from almost maximum degree sized palettes}
\author{Landon Rabern}
\date{\today}

\begin{document}
\maketitle

\section{Introduction}
My dissertation contains material on a few different topics all relating to graph coloring, but today I'll mostly talk about a conjecture of Borodin and Kostochka from 1977.  But before i get to that, i want to put the problem in context.  First, i need to define some terms.

Define, $\Delta(G)$, $K_t$, coloring.  PICTURES.

\bigskip

A pair of complimentary results from the late 90s tell us that $c$-coloring transitions from being hard to being easy around around $c = \Delta - \sqrt{\Delta}$.  The first is from Emden-Weinert, Hougardy and Kreuter and says:

\begin{thm}
Fix $\Delta$.  For any $k$ such that $k^2 + k > \Delta$, the problem of determining whether a graph $G$ of maximum degree $\Delta$ has a $(\Delta + 1 - k)$-coloring is NP-complete (also need, $\Delta + 1 - k \ge 3$).
\end{thm}

The second result is from Molloy and Reed.

\begin{thm}
There exists $\Delta_0$ such that for fixed $\Delta \geq \Delta_0$ and $k$ such that $k^2 + k \leq \Delta$ the problem of determining whether a graph $G$ of maximum degree $\Delta$ has a $(\Delta + 1 - k)$-coloring is in P.
\end{thm}

The large $\Delta$ requirement comes from using the probabilistic method.  The complexity situation for small $\Delta$ is open.  For $k = 0$, this was solved in 1941 by Brooks:

\begin{thm}[Brooks 1941]
Every graph $G$ with $\Delta(G) \geq 3$ that doesn't contain $K_{\Delta(G) + 1}$ is $\Delta(G)$-colorable.
\end{thm} 

So, an algorithm can just test for a $K_{\Delta + 1}$ component.  In 1977, Borodin and Kostochka conjectured a result that would solve the case when $k = 1$ and $\Delta \geq 9$.

\begin{conjecture}
Every graph $G$ with $\Delta(G) \geq 9$ that doesn't contain $K_{\Delta(G)}$ is $(\Delta(G)-1)$-colorable.
\end{conjecture}

Talk, $K_\Delta$ is the obvious obstruction to $(\Delta-1)$-coloring.  The $\Delta \geq 9$ condition is necessary:

\input{SmallCE}

\bigskip
\noindent Known results:
\begin{itemize}
\item In 1980, Kostochka proved that if we exclude $K_{\Delta(G)-29}$ instead, then $G$ is $(\Delta(G)-1)$-colorable.
\item Later in the 1980s, Mozhan proved that if $\Delta(G) \geq 31$ and we exclude $K_{\Delta(G)-3}$ instead, then $G$ is $(\Delta(G)-1)$-colorable.
\item In 1999, Reed proved that the conjecture holds for $\Delta \geq 10^{14}$.  
\end{itemize}

\noindent Highlights:
\begin{itemize}
\item We prove the full Borodin-Kostochka conjecture for claw-free graphs.
\item We prove that the following conjecture is equivalent to the Borodin-Kostochka conjecture.
\begin{conjecture}
If $G$ is a graph with $\Delta(G) \geq 9$ such that $G$ doesn't contain $\join{K_3}{\overline{K}_{\Delta(G)-3}}$, then $G$ is $(\Delta(G)-1)$-colorable.
\end{conjecture}
\item We generalize Reed's result to list coloring:

\begin{thm}
There exists $\Delta_0$ such that every graph with $\Delta \geq \Delta_0$ that doesn't contain $K_{\Delta}$ is $(\Delta-1)$-choosable.
\end{thm}

Define $k$-choosable.
\end{itemize}

\section{List coloring lemmas}
We call a graph $d_k$-choosable if it colorable from any list assignment $L$ with $\card{L(v)} \geq d(v) - k$ for each vertex $v$.  Why do we care about these particular list assignments?  Suppose we want to $(\Delta(G) - k)$-color a graph $G$ and that we can $(\Delta(G) - k)$-color some proper induced subgraph $H$.  PICTURE.  Look at the lists of colors available on $G-H$.  Each $v \in V(G-H)$ has at most $\Delta(G) - d_{G-H}(v)$ neighbors in $H$, so has at least $\Delta(G) - k - (\Delta(G) - d_{G-H}(v)) = d_{G-H}(v) - k$ colors available.  So, if $G-H$ is $d_k$-choosable, we can finish.

The $d_0$-choosable graphs were classified in the 70s by Borodin and independently Erdos-Rubin-Taylor.  For the Borodin-Kostochka conjecture, we want to know about $d_1$-choosable graphs.

We classified the $d_1$-choosable graph joins $\join{A}{B}$ where $\card{A}, \card{B} \geq 2$, this gives a lot of structure about a counterexample to BK.  The classification takes 45 pages to prove, we won't go into it, but will use the results to prove Borodin-Kostochka for claw-free graphs.  One example, $\join{K_6}{B}$ is $d_1$-choosable unless $\omega(B) \geq \card{B} - 1$, so intersections of cliques are severely restricted.  PICTURE.

\section{Borodin-Kostochka for claw-free graphs}

\bigskip
We outline the proof of the following.
\begin{thm}
Every claw-free graph with $\Delta \geq 9$ that doesn't contain $K_{\Delta}$ can be $(\Delta-1)$-colored.
\end{thm}

The proof uses the structure theorem for claw-free graphs proved by Chudnovsky and Seymour. We actually only need a simpler part of it: the structure theorem for quasi-line graphs; graphs where the neighborhood of every vertex can be covered by two cliques.  PICTURE.

\bigskip

We use the following structure theorem for quasi-line graphs.

\begin{lem}\label{QuasilineStructure}
Every connected skeletal quasi-line graph is a circular interval graph or a composition of
linear interval strips.
\end{lem}

We need to define the terms in this lemma.

A \emph{homogeneous pair of cliques} $(A_1, A_2)$ in a graph $G$ is a pair of
disjoint nonempty cliques such that for each $i \in \irange{2}$, every vertex in
$G - (A_1 \cup A_2)$ is either joined to $A_i$ or misses all of $A_i$ and
$\card{A_1} + \card{A_2} \geq 3$. PICTURES.

A homogeneous pair of cliques $(A_1, A_2)$ is $\emph{skeletal}$
if for any $e \in E(A, B)$ we have $\omega(G[A \cup B] - e) < \omega(G[A \cup
B])$.  A graph is $\emph{skeletal}$ if it contains no nonskeletal homogeneous
pair of cliques.  

\bigskip

Given a set $V$ of points on the unit circle together with a set of closed intervals $C$ on the unit circle we define a graph with vertex set $V$ and an edge between two different vertices if and only if they are both contained in some element of $C$.  Any graph isomorphic to such a graph is a \emph{circular interval graph}.  Similarly, by replacing the unit circle with the unit interval, we get the class of \emph{linear interval graphs}.

\bigskip

It remains to define \emph{compositions of linear interval strips}.  These are a generalization of line graphs. 
A \emph{linear interval strip} $(S, A_1, A_2)$ is a linear interval graph $S$ together with end cliques $A_1$ and $A_2$.  PICTURE.

\bigskip

Let $H$ be a directed multigraph (possibly with loops)
and suppose for each edge $e$ of $H$ we have a strip $(S_e, X_e, Y_e)$.  For
each $v \in V(H)$ define

\[C_v \DefinedAs \parens{\bigcup \setbs{X_e}{\text{$e$ is directed out of $v$}}}
\cup \parens{\bigcup \setbs{Y_e}{\text{$e$ is directed into $v$}}}\]

The graph formed by taking the disjoint union of $\setbs{S_e}{e \in E(H)}$ and
making $C_v$ a clique for each $v \in V(H)$ is the composition of the strips
$(S_e, X_e, Y_e)$.  Any graph formed in such a manner is called a
\emph{composition of linear interval strips}.  PICTURE.

Taking all strips to have a single vertex gives the line graph construction.

\bigskip\bigskip

Now we can outline the proof.

\begin{enumerate}
\item Prove for circular interval graphs.
\item Reduce from quasi-line graphs to line graphs as follows:
	\begin{enumerate}
	\item It is always possible to make skeletal counterexample from a given counterexample just by removing edges in nonskeletal homogeneous pairs of cliques.  Do so.
	\item We must have a composition of linear interval strips by the structure theorem.
	\item Take a composition representation using the maximum number of strips.
	\item Show that for each strip $(S, A_1, A_2)$ we must have $V(S) = A_1 = A_2$ and thus we have a line graph.
	\end{enumerate}
\item Prove for line graphs of multigraphs.
\item Reduce from claw-free graphs to quasi-line graphs.
\end{enumerate}

Steps (1), (2) and (4) all rely heavily on our classification of $d_1$-choosable joins.  Step (4) uses some $d_1$-choosability results outside this classification, for example, the following graph $D_8$ is $d_1$-choosable:

\input{D8}

The reduction from claw-free graphs to quasi-line graphs works for list coloring as well.  Also, the circular interval graphs proof works for list coloring.  So, the following generalization seems within reach.

\begin{conjecture}
Every claw-free graph with $\Delta \geq 9$ that doesn't contain $K_{\Delta}$ is $(\Delta-1)$-choosable.
\end{conjecture}

Borodin and Kostochka conjecture that this holds with the claw-free restriction removed.  As evidence of this, we generalized Reed's proof of Borodin-Kostochka for large $\Delta$ to list coloring, proving:

\begin{thm}
There exists $\Delta_0$ such that every graph with $\Delta \geq \Delta_0$ that doesn't contain $K_{\Delta}$ is $(\Delta-1)$-choosable.
\end{thm}

\section{Future directions}
\begin{enumerate}
\item BK for list coloring for claw-free graphs.
\item improved bounds on the number of edges in (online) list-critical.  These can be used to prove Ore degree bounds for (online) list coloring; we have so far that the Ore degree version of Brooks' theorem for (online) list coloring holds for $\Delta \geq 11$.
\item Improve Mozhan's methods to get down to $\Delta-2$, we can now prove his $\Delta-3$ result for $\Delta \geq 13$ instead of $\Delta \geq 31$, and the proof is relatively simple.
\item BK for large $\Delta$ for online list coloring (and Alon-Tarsi number)
\end{enumerate}
\end{document}
