\documentclass[12pt]{article}

\title{}
\begin{document}
\maketitle
Trying to synthesize all the new incoming ideas overwhelms sometimes, feels like there is this machine i can lift my brain up into and see all the things
at the same time. but hard to stay there, but i think that is fine, in and out, knowing what it is and that it is not bad give me control. i was thinking
of people's use of the phrase 'psychotic break'.  i think i see what that is now, it is breaking through, seeing and getting scared.  but meditation appears
to be able to meta the fear.  then it is just seeing farther, meta the fear which may itself be scary and meta that and so on.  write when i feel
like writing, this log instead of just typing it to Bobby.

i meditated myself into a state where my parts felt plastic, i could move if i wanted, but i didn't.  then it morphed into a state
i have seen before only briefly.  full deep meditation, eyes closed but using them like they are open.  i tried moving my hands and
i could do it normally, actually everything normally, just that i was blind.  i felt over to the edge of the bed and put
my feet on the ground to see if i could stand up.  i could, just normally, so i felt around a bit for the wall and the door
opening.  i felt my way to a nearby chair and sat down wondering how long i could maintain this everything normal but trading vision for a sort
of meditation view of things.  like i am looking around as if my eyes were open trying to see things, i do see outlines of things morphing in
and out, i think sometimes when i know where something is like my hands, my brain is telling me so how with ripples in this weird visual field.
anyway, i tried for a bit to warp in objects that i had seen before, into the visual field so i could do something with them.  i managed to get
a nice blue color, but then i had to pee.  so, i felt my way over to the bathroom and did that and came back.  at this point the meditation depth
was starting to decrease, but still positive.  i wanted to know what time it was since the baby alf was down stairs with granny and atticus and
i didn't want to put her out too much.  but i didn't want to end the state.  i had a funny idea, felt around for my phone, said "ok google" at it
a bunch of times to no effect since my speech was a little off.  managed to get the speach clear and google vibrated back, i asked "what time is it?"
but got no reply.  media sound was off, after many iterations of tapping somewhere on the phone and then trying "ok google" again, i finally got an audible
answer to "what is a donkey?" then i asked "what time is it?" and google told me 3:41pm.  i smiled at the victory and opened my eyes, back to normal sight.

i just tried really hard to levitate.  thinking along these lines, if zombies are conceivable, then possible, so there must be some sort of consciousness particles
or something like that we are misssing.  whatever that is can be harnassed via meditation to produce physical effects.  those are the assumptions i am trying
out now.  i really thought i might levitate, all body parts plasticified and my legs feeling more and more pressure as i attempted to "push" off
those consciousness particles i am assuming, it started to hurt like weight lifting kind of fatigue, pushed more since probably there should be some
pain if i am going to fly, got to a point where i could keep going and i knew it would hurt much more, at this point i am not sure what exactly is happening,
one likely hypothesis is that my body is trying as hard as it possibly can to do an impossible task, so i think i might break something if i
continue, so i backed off.  will try again and see if practice helps.

\end{document}

