\documentclass[oneside,12pt]{memoir}

%%% Garamond
%\usepackage[T1]{fontenc}
%\usepackage{lmodern}
%\usepackage{garamond}

\DoubleSpacing         % memoir's double spacing
\usepackage{pwasu}     % this package
\usepackage[paperwidth=8.5in,paperheight=11.0in,
  left=1.25in,right=1.25in,top=1.0in,bottom=1.0in,
  includefoot]{geometry}

\usepackage{amssymb, amsmath, amsthm}

\usepackage{tkz-graph}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}
\usepackage[position=bottom]{subfig}

\usepackage{longtable}
\usepackage{array}

%\usepackage[pdftex, colorlinks, hyperfootnotes]{hyperref}

\makeatletter
\newtheorem*{rep@theorem}{\rep@title}
\newcommand{\newreptheorem}[2]{
\newenvironment{rep#1}[1]{
 \def\rep@title{#2 \ref{##1}}
 \begin{rep@theorem}}
 {\end{rep@theorem}}}
\makeatother


\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newreptheorem{thm}{Theorem}
\newtheorem{prop}[thm]{Proposition}
\newreptheorem{prop}{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newreptheorem{lem}{Lemma}
\newtheorem{conjecture}[thm]{Conjecture}
\newreptheorem{conjecture}{Conjecture}
\newtheorem{cor}[thm]{Corollary}
\newreptheorem{cor}{Corollary}
\newtheorem{prob}[thm]{Problem}
\newtheorem{claim}{Claim}
\newtheorem*{unnumberedClaim}{Claim}
\newtheorem*{SmallPotLemma}{Small Pot Lemma}
\newtheorem*{BK}{Borodin-Kostochka Conjecture}
\newtheorem*{BK2}{Borodin-Kostochka Conjecture (restated)}
\newtheorem*{Reed}{Reed's Conjecture}
\newtheorem*{ClassificationOfd0}{Classification of $d_0$-choosable graphs}

\theoremstyle{definition}
\newtheorem{defn}{Definition}[section]
\newtheorem*{CliqueGraph}{Clique Graph}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem*{question}{Question}
\newtheorem*{observation}{Observation}

\newcommand{\fancy}[1]{\mathcal{#1}}
\newcommand{\C}[1]{\fancy{C}_{#1}}
\newcommand{\IN}{\mathbb{N}}
\newcommand{\IR}{\mathbb{R}}
\newcommand{\G}{\fancy{G}}
\newcommand{\CC}{\fancy{C}}
\newcommand{\D}{\fancy{D}}

\newcommand{\inj}{\hookrightarrow}
\newcommand{\surj}{\twoheadrightarrow}

\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\setb}[3]{\left\{ #1 \in #2 \mid #3 \right\}}
\newcommand{\setbs}[2]{\left\{ #1 \mid #2 \right\}}
\newcommand{\card}[1]{\left|#1\right|}
\newcommand{\size}[1]{\left\Vert#1\right\Vert}
\newcommand{\ceil}[1]{\left\lceil#1\right\rceil}
\newcommand{\floor}[1]{\left\lfloor#1\right\rfloor}
\newcommand{\func}[3]{#1\colon #2 \rightarrow #3}
\newcommand{\funcinj}[3]{#1\colon #2 \inj #3}
\newcommand{\funcsurj}[3]{#1\colon #2 \surj #3}
\newcommand{\irange}[1]{\left[#1\right]}
\newcommand{\join}[2]{#1 \mbox{\hspace{2 pt}$\ast$\hspace{2 pt}} #2}
\newcommand{\djunion}[2]{#1 \mbox{\hspace{2 pt}$+$\hspace{2 pt}} #2}
\newcommand{\parens}[1]{\left( #1 \right)}
\newcommand{\brackets}[1]{\left[ #1 \right]}

\newcommand{\DefinedAs}{\mathrel{\mathop:}=}

\newcommand{\mov}[2]{#1^{#2}}
\newcommand{\wt}[1]{w\parens{#1}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\def\adj{\leftrightarrow}
\def\nonadj{\not\!\leftrightarrow}
\newcommand{\im}{\operatorname{im}}
\newcommand{\ex}{\operatorname{E}}

\newcommand{\texorpdfstring}[2]{#1}

% any changes to \claim should be mirrored in \claimnonum and \subclaim
\renewcommand{\claim}[2]{\textbf{Claim #1.}~\textit{#2}~~}
\newcommand{\claimnonum}[1]{\textbf{Claim.}~\textit{#1}~~}
\newcommand{\subclaim}[2]{\textbf{Subclaim #1.}~\textit{#2}~~}

\setcounter{tocdepth}{3}
\maxsecnumdepth{subsection}

%\%\topmargin -0.4in        % read Lamport p.163
%\topmargin -0.36in        % read Lamport p.163
%\oddsidemargin 0.5in   % read Lamport p.163
%\evensidemargin 0.5in  	% same as oddsidemargin but for left-hand pages
%\textwidth 6in
%\usepackage[textheight = 9.6in]{geometry}
%\textheight 9.03in 
%\pagestyle{empty}       % Uncomment if you don't want page numbers
\parskip 7.2pt           % sets spacing between paragraphs
%\renewcommand{\baselinestretch}{1.5} % Uncomment for 1.5 spacing between lines
\parindent 0pt          % sets leading space for paragraphs
 %%%%%%%%

\DoubleSpacing
\begin{document}
\maxtocdepth{subparagraph}
\pagestyle{plain}
\frontmatter
\thetitlepage

\asuabstract
\setcounter{page}{1}
\setlength{\parindent}{.5in} 
Every graph can be colored with one more color than its maximum degree. A
well-known theorem of Brooks gives the precise conditions under which a graph
can be colored with maximum degree colors.  It is natural to ask for the
required conditions on a graph to color with one less color than the maximum
degree; in 1977 Borodin and Kostochka conjectured a solution for graphs with maximum degree at least 9: as long as the
graph doesn't contain a maximum-degree-sized clique, it can be colored with one
fewer than the maximum degree colors.

This study attacks the conjecture on multiple fronts.  The first technique is an
extension of a vertex shuffling procedure of Catlin and is used to prove the
conjecture for graphs with edgeless high vertex subgraphs.  This general
approach also bears more theoretical fruit.

The second technique is an extension of a method Kostochka used to reduce the
Borodin-Kostochka conjecture to the maximum degree 9 case.  Results on the
existence of independent transversals are used to find an independent set
intersecting every maximum clique in a graph.  

The third technique uses list coloring results to exclude induced subgraphs in a
counterexample to the conjecture. The classification of such excludable graphs
that decompose as the join of two graphs is the backbone of
many of the results presented here.

The fourth technique uses the structure theorem for quasi-line graphs of 
Chudnovsky and Seymour in concert with the third technique to prove the
Borodin-Kostochka conjecture for claw-free graphs.
\newpage
The fifth technique adds edges to proper induced subgraphs of a minimum
counterexample to gain control over the colorings produced by minimality.

The sixth technique adapts a recoloring technique originally developed for
strong coloring by Haxell and by Aharoni, Berger and Ziv to general coloring. 
Using this recoloring technique, the Borodin-Kostochka conjectured is proved for
graphs where every vertex is in a large clique.

The final technique is naive probabilistic coloring as employed by
Reed in the proof of the Borodin-Kostochka conjecture for large maximum degree. 
The technique is adapted to prove the Borodin-Kostochka conjecture for list
coloring for large maximum degree.

\newgeometry{paperwidth=8.5in,paperheight=11.0in,
  left=1.25in,right=1.25in,top=1.1in,bottom=1.0in,
  includefoot}
\vspace*{-0.13in}
\tableofcontents*
\restoregeometry

\listoffigures

\mainmatter
\pagestyle{asu}

\addcontentsline{toc}{chapter}{CHAPTER}
\pagestyle{plain}
\setlength{\parindent}{.5in}
\input{History}
\input{Brooks}
\input{VertexShuffle}
\input{Reduce}
\input{ListColoring}
\input{ClawFree}
\input{mules}
\input{StrongColoring}
\input{ChoiceLargeDelta}
\newpage
\begingroup
\let\clearpage\relax
\vspace*{0.22in}
\SingleSpace
\begin{thebibliography}{99}
\bibitem{aharoni2007independent}
R.~Aharoni, E.~Berger, and R.~Ziv, \emph{Independent systems of representatives
  in weighted graphs}, Combinatorica \textbf{27} (2007), no.~3, 253--267.

\bibitem{Alon19997}
N.~Alon, \emph{Combinatorial nullstellensatz}, Combinatorics Probability and
  Computing \textbf{8} (1999), no.~1--2, 7--29.

\bibitem{Alon1992125}
N.~Alon and M.~Tarsi, \emph{Colorings and orientations of graphs},
  Combinatorica \textbf{12} (1992), no.~2, 125--134.

\bibitem{anderson1977edge}
L.D. Andersen, \emph{On edge-colorings of graphs}, Math. Scand \textbf{40}
  (1977), 161--175.

\bibitem{beutelspacher1984minimal}
A.~Beutelspacher and P.R. Hering, \emph{Minimal graphs for which the chromatic
  number equals the maximal degree}, Ars Combin \textbf{18} (1984), 201--216.

\bibitem{bollobasManvel}
B.~Bollob{\'a}s and B.~Manvel, \emph{Optimal vertex partitions}, Bulletin of
  the London Mathematical Society \textbf{11} (1979), no.~2, 113.

\bibitem{BorodinKostochkaWoodall}
O.V. Borodin, A.V. Kostochka, and D.R. Woodall, \emph{List edge and list
  total colourings of multigraphs}, J. Combin. Theory Ser. B \textbf{71}
  (1997), no.~2, 184--204.

\bibitem{borodin1976decomposition}
O.V. Borodin, \emph{On decomposition of graphs into degenerate subgraphs},
  Metody Diskretn. Analiz \textbf{28} (1976), 3--11 (in Russian).

\bibitem{borodin1977criterion}
O.V. Borodin, \emph{{Criterion of chromaticity of a degree prescription}}, Abstracts
  of IV All-Union Conf. on Th. Cybernetics, 1977, pp.~127--128.

\bibitem{borodin1977upper}
O.V. Borodin and A.V. Kostochka, \emph{{On an upper bound of a graph's
  chromatic number, depending on the graph's degree and density}}, Journal of
  Combinatorial Theory, Series B \textbf{23} (1977), no.~2-3, 247--250.

\bibitem{borodin2000variable}
O.V. Borodin, A.V. Kostochka, and B.~Toft, \emph{{Variable degeneracy:
  extensions of Brooks' and Gallai's theorems}}, Discrete Mathematics
  \textbf{214} (2000), no.~1-3, 101--112.

\bibitem{brooks1941colouring}
R.L. Brooks, \emph{{On colouring the nodes of a network}}, Mathematical
  Proceedings of the Cambridge Philosophical Society, vol.~37, Cambridge Univ
  Press, 1941, pp.~194--197.

\bibitem{cariolaro2006fans}
D.~Cariolaro, \emph{On fans in multigraphs}, Journal of Graph Theory
  \textbf{51} (2006), no.~4, 301--318.

\bibitem{CatlinAnotherBound}
P.A. Catlin, \emph{{Another bound on the chromatic number of a graph}},
  Discrete Mathematics \textbf{24} (1978), no.~1, 1--6.

\bibitem{catlin1978bound}
P.A. Catlin, \emph{A bound on the chromatic number of a graph}, Discrete
  Mathematics \textbf{22} (1978), no.~1, 81--83.

\bibitem{catlin1979hajos}
P.A. Catlin, \emph{Haj{\'o}s’ graph-coloring conjecture: variations and
  counterexamples}, J. Combin. Theory Ser. B \textbf{26} (1979), no.~2,
  268--274.

\bibitem{chudnovskyFradkin}
M.~Chudnovsky and A.~Ovetsky, \emph{Coloring quasi-line graphs}, Journal of
  Graph Theory \textbf{54} (2007), no.~1, 41--50.

\bibitem{chudnovsky2005structure}
M.~Chudnovsky and P.~Seymour, \emph{The structure of claw-free graphs}, Surveys
  in combinatorics \textbf{327} (2005), 153--171.

\bibitem{cranstonrabernclaw}
D.W. Cranston and L.~Rabern, \emph{{Coloring claw-free graphs with $\Delta-1$
  colors}}, SIAM J. Discrete Math. (Forthcoming).

\bibitem{mules}
D.W. Cranston and L.~Rabern, \emph{{Conjectures equivalent to the Borodin-Kostochka Conjecture that
  appear weaker}}, arXiv:1203.5380 (2012).

\bibitem{dhurandhar1982improvement}
M.~Dhurandhar, \emph{Improvement on {Brooks}' chromatic bound for a class of
  graphs}, Discrete Mathematics \textbf{42} (1982), no.~1, 51--56.

\bibitem{diestel2010}
R.~Diestel, \emph{{Graph Theory}}, 4 ed., Springer-Verlag, Heidelberg, 2010.

\bibitem{Entringer1985367}
R.C. Entringer, \emph{{A Short Proof of Rubin's Block Theorem}}, {Annals of
  Discrete Mathematics 27 - Cycles in Graphs} (B.R. Alspach and C.D. Godsil,
  eds.), North-Holland Mathematics Studies, vol. 115, North-Holland, 1985,
  pp.~367--368.

\bibitem{erdos1979choosability}
P.~Erd\H{o}s, A.L. Rubin, and H.~Taylor, \emph{{Choosability in graphs}}, Proc.
  West Coast Conf. on Combinatorics, Graph Theory and Computing, Congressus
  Numerantium, vol.~26, 1979, pp.~125--157.

\bibitem{StiebitzVizingGoldberg}
L.M. Favrholdt, M.~Stiebitz, and B.~Toft, \emph{Graph edge colouring: Vizing's
  theorem and goldberg's conjecture},  (2006).

\bibitem{gallai1963kritische}
T.~Gallai, \emph{Kritische graphen i.}, Math. Inst. Hungar. Acad. Sci
  \textbf{8} (1963), 165--192 (in German).

\bibitem{GoldbergJGT}
M.K. Goldberg, \emph{Edge-coloring of multigraphs: Recoloring technique},
  Journal of Graph Theory \textbf{8} (1984), no.~1, 123--137.

\bibitem{gravier1998graphs}
S.~Gravier and F.~Maffray, \emph{Graphs whose choice number is equal to their
  chromatic number}, Journal of Graph Theory \textbf{27} (1998), no.~2, 87--97.

\bibitem{grötschel1981ellipsoid}
M.~Gr\"{o}tschel, L.~Lov\'{a}sz, and A.~Schrijver, \emph{{The ellipsoid method
  and its consequences in combinatorial optimization}}, Combinatorica
  \textbf{1} (1981), no.~2, 169--197.

\bibitem{HajnalSaturation}
A.~Hajnal, \emph{{A theorem on $k$-saturated graphs}}, Canadian Journal of
  Mathematics \textbf{17} (1965), no.~5, 720.

\bibitem{haxell2001note}
P.~Haxell, \emph{A note on vertex list colouring}, Combinatorics, Probability
  and Computing \textbf{10} (2001), no.~04, 345--347.

\bibitem{haxell2004strong}
P.~Haxell, \emph{On the strong chromatic number}, Combinatorics, Probability
  and Computing \textbf{13} (2004), no.~06, 857--865.

\bibitem{haxell2011forming}
P.~Haxell, \emph{On forming committees}, The American Mathematical Monthly
  \textbf{118} (2011), no.~9, 777--788.

\bibitem{haxell2006odd}
P.~Haxell and T.~Szab{\'o}, \emph{Odd independent transversals are odd},
  Combinatorics Probability and Computing \textbf{15} (2006), no.~1/2, 193.

\bibitem{Hladky}
J.~Hladk{\`y}, D.~Kr{\'a}l, and U.~Schauz, \emph{{Brooks' Theorem via the
  Alon-Tarsi Theorem}}, Discrete Mathematics (2010).

\bibitem{GraphColoringProblems}
T.R. Jensen and B.~Toft, \emph{Graph coloring problems}, John Wiley \& Sons,
  1995.

\bibitem{kierstead2000choosability}
H.A. Kierstead, \emph{{On the choosability of complete multipartite graphs with
  part size three}}, Discrete Mathematics \textbf{211} (2000), no.~1-3,
  255--259.

\bibitem{kierstead2009ore}
H.A. Kierstead and A.V. Kostochka, \emph{{Ore-type versions of Brooks'
  theorem}}, Journal of Combinatorial Theory, Series B \textbf{99} (2009),
  no.~2, 298--305.

\bibitem{kierstead1986chromatic}
H.A. Kierstead and J.H. Schmerl, \emph{The chromatic number of graphs which
  induce neither {$K_{1,3}$} nor {$K_5-e$}}, Discrete mathematics \textbf{58}
  (1986), no.~3, 253--262.

\bibitem{KingPersonalKostochkaGeneralized}
A.~King, personal communication, 2009.

\bibitem{king2009claw}
A. King, \emph{{Claw-free graphs and two conjectures on omega, Delta, and
  chi}}, Ph.D. thesis, McGill University, 2009.

\bibitem{KingHitting}
A. King, \emph{Hitting all maximum cliques with a stable set using lopsided
  independent transversals}, Journal of Graph Theory (2010).

\bibitem{king2008bounding}
A. King and B.A. Reed, \emph{{Bounding $\chi$ in terms of $\omega$ and
  $\Delta$ for quasi-line graphs}}, Journal of Graph Theory \textbf{59} (2008),
  no.~3, 215--228.

\bibitem{king2007upper}
A. King, B.A. Reed, and A.~Vetta, \emph{{An upper bound for the chromatic
  number of line graphs}}, European Journal of Combinatorics \textbf{28}
  (2007), no.~8, 2182--2187.

\bibitem{kostochkaRussian}
A.V. Kostochka, \emph{{Degree, density, and chromatic number}}, Metody
  Diskretn. Analiz \textbf{35} (1980), 45--70 (in Russian).

\bibitem{KostochkaTriangleFree}
A.V. Kostochka, \emph{{A modification of Catlin's algorithm}}, Methods and Programs of
  Solutions Optimization Problems on Graphs and Networks (1982), no.~2, 75--79
  (in Russian).

\bibitem{PersonalComms}
A.V. Kostochka, personal communication, 2012.

\bibitem{krs_one}
A.V. Kostochka, L.~Rabern, and M.~Stiebitz, \emph{{Graphs with chromatic number
  close to maximum degree}}, Discrete Mathematics \textbf{312} (2012), no.~6,
  1273--1281.

\bibitem{BrooksExtended}
A.V. Kostochka, M.~Stiebitz, and B.~Wirth, \emph{{The colour theorems of Brooks
  and Gallai extended}}, Discrete Mathematics \textbf{162} (1996), no.~1-3,
  299--303.

\bibitem{lawrence1978covering}
J.~Lawrence, \emph{Covering the vertex set of a graph with subgraphs of smaller
  degree}, Discrete Mathematics \textbf{21} (1978), no.~1, 61--68.

\bibitem{lovasz1966decomposition}
L.~Lov\'{a}sz, \emph{On decomposition of graphs}, Studia Sci. Math. Hungar.
  \textbf{1} (1966), 237--238.

\bibitem{Lovasz1975269}
L. Lov\'{a}sz, \emph{Three short proofs in graph theory}, Journal of Combinatorial
  Theory, Series B \textbf{19} (1975), no.~3, 269--271.

\bibitem{mader}
W.~Mader, \emph{{Existenz n-fach zusammenh{\"a}ngender Teilgraphen in Graphen
  gen{\"u}gend gro{\ss}er Kantendichte}}, {Abhandlungen aus dem Mathematischen
  Seminar der Universit{\"a}t Hamburg}, vol.~37, Springer, 1972, pp.~86--97.

\bibitem{molloy2002graph}
M.S. Molloy and B.A. Reed, \emph{{Graph colouring and the probabilistic
  method}}, Springer Verlag, 2002.

\bibitem{mozhan1983}
N.N. Mozhan, \emph{{Chromatic number of graphs with a density that does not
  exceed two-thirds of the maximal degree}}, Metody Diskretn. Analiz
  \textbf{39} (1983), 52--65 (in Russian).

\bibitem{naserasr}
R.~Naserasr, personal communication, 2008.

\bibitem{niessen2000round}
T.~Niessen and J.~Kind, \emph{The round-up property of the fractional chromatic
  number for proper circular arc graphs}, Journal of Graph Theory \textbf{33}
  (2000), no.~4, 256--267.

\bibitem{2009arXiv0907.3705R}
L.~Rabern, \emph{{On hitting all maximum cliques with an independent set}},
  Arxiv preprint arXiv:0907.3705 (2009).

\bibitem{rabernhitting}
L. Rabern, \emph{{On hitting all maximum cliques with an independent set}},
  Journal of Graph Theory \textbf{66} (2011), no.~1, 32--37.

\bibitem{rabern2010b}
L. Rabern, \emph{{An improvement on Brooks' theorem}}, Manuscript (2010).

\bibitem{rabern2011strengthening}
L. Rabern, \emph{{A strengthening of Brooks' Theorem for line graphs}}, Electron.
  J. Combin. \textbf{18} (2011), no.~p145, 1.

\bibitem{partitionnote}
L. Rabern, \emph{A note on vertex partitions}, Manuscript (2011).

\bibitem{rabern2010a}
L. Rabern, \emph{{$\Delta$-critical graphs with small high vertex cliques}},
  Journal of Combinatorial Theory, Series B \textbf{102} (2012), no.~1,
  126--130.

\bibitem{rabern2010destroying}
L. Rabern, \emph{Destroying non-complete regular components in graph partitions},
  J. Graph Theory (Forthcoming).

\bibitem{rabern2012partitioning}
L. Rabern, \emph{Partitioning and coloring with degree constraints}, Discrete
  Math. (Forthcoming).

\bibitem{reed1998omega}
B.~Reed, \emph{{$\omega$, $\Delta$, and $\chi$}}, Journal of Graph Theory
  \textbf{27} (1998), no.~4, 177--212.

\bibitem{reed1999strengthening}
B. Reed, \emph{{A strengthening of Brooks' theorem}}, Journal of Combinatorial
  Theory, Series B \textbf{76} (1999), no.~2, 136--149.

\bibitem{ReedSudakov}
B.~Reed and B.~Sudakov, \emph{{List colouring when the chromatic number is
  close to the order of the graph}}, Combinatorica \textbf{25} (2004), no.~1,
  117--123.

\bibitem{schauz2009mr}
U.~Schauz, \emph{{Mr. Paint and Mrs. Correct}}, {The Electronic Journal of
  Combinatorics} \textbf{16} (2009), no.~1, R77.

\bibitem{stacho2001new}
L.~Stacho, \emph{{New upper bounds for the chromatic number of a graph}},
  Journal of Graph Theory \textbf{36} (2001), no.~2, 117--120.

\bibitem{stiebitz1982proof}
M.~Stiebitz, \emph{{Proof of a conjecture of T. Gallai concerning connectivity
  properties of colour-critical graphs}}, Combinatorica \textbf{2} (1982),
  no.~3, 315--323.

\bibitem{szabo2006extremal}
T.~Szab{\'o} and G.~Tardos, \emph{Extremal problems for transversals in graphs
  with bounded degree}, Combinatorica \textbf{26} (2006), no.~3, 333--351.

\bibitem{tverberg1983brooks}
H.~Tverberg, \emph{{On Brooks' theorem and some related results}}, Mathematics
  Scandinavia \textbf{52} (1983), 37--40.
\end{thebibliography}

\endgroup

\addcontentsline{toc}{chapter}{APPENDIX}
\clearpage
\appendix
\chapter{NOTATION}\label{NotationAppendix}
\newpage
\begin{Spacing}{1}
\vspace*{-0.5in}
\input{Notation}
\end{Spacing}
\end{document}

